Performance Evaluation of Big Data Processing Strategies for Neuroimaging

### 2018-12-12 paper.tex
Results section and fixes to figures

\usepackage{graphicx}
\usepackage{color}
\usepackage{colortbl} % for \rowcolor

\usepackage[bookmarks=false]{hyperref}
\usepackage{algpseudocode}
\usepackage{algorithm}
\algnewcommand\algorithmicforeach{\textbf{for each}}
\algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}

@@

@@

\subsection{CPU Time}
% Comment on the general trend
% We can't really say that imaging tasks are not small. Even Freesurfer could be decomposed.
%

\subsection{Image Block size}
% Comment on the general trend

\subsection{Data Size}
% Comment on the general trend

\subsection{Iterations}
% Comment on the general trend

>>>>

\usepackage{graphicx}
\usepackage{color}
\usepackage{colortbl} % for \rowcolor
\usepackage{caption}
\usepackage{subcaption}
\usepackage[bookmarks=false]{hyperref}
\usepackage{algpseudocode}
\usepackage{algorithm}
\algnewcommand\algorithmicforeach{\textbf{for each}}
\algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}

@@

\begin{figure*}
\begin{subfigure}{0.33\linewidth}
\centering
    \includegraphics[width=\textwidth]{results/figures/tmpfs-incrementation.pdf}
\caption{}
\label{fig:sub1}
\end{subfigure}%
    \begin{subfigure}{0.33\linewidth}
\centering
    \includegraphics[width=\textwidth]{results/figures/local-incrementation.pdf}
\caption{}
\label{fig:sub2}
    \end{subfigure}%
\begin{subfigure}{0.33\linewidth}
\centering
    \includegraphics[width=\textwidth]{results/figures/lustre-incrementation.pdf}
\caption{}
\label{fig:sub3}
\end{subfigure}
    \caption{Model evaluation on (a) tmpfs, (b) local and (c) Lustre. Grey 
             regions denote areas that violate model predictions.}
\label{fig:modeleval}
\end{figure*}

@@

In order to evaluate our model, we compared the observed in-memory speedup 
ratio to our model's expected speed-up ratio for the given 
D/C / $\delta$/$\gamma$ used (Figure~\ref{fig:modeleval}). Experiments for 
which there was no in-memory equivalent (i.e. BigBrain split into 30 chunks) 
were not considered.

%probably confusing to mention values, but they're here for now
Results show that, overall, the model correctly predicted the effect of page 
cache on processing times for local disk and Lustre. Points which violated 
model predictions were found at 1 iteration, where page cache would not have 
been saturated, due to the amount of data being written, and at a task duration
of 320 seconds. However, in all cases, the ``1" boundary was never trespassed by 
more than a factor of 0.19, and is therefore likely a result of system 
variability.

Lustre's regression slope is smallest of the three filesystems, possibly 
lending to the fact that Lustre was nearly always in transient state. Out of 
the thirty four benchmarks taken on Lustre, Lustre was only expected to be in steady 
state for four of the benchmarks (MRI and all benchmarks with a steady state of
320). This was validated by our experiments.

The next smallest slope is that of the local disk. The local disk was \todo{continue after disk benchmarks}

Unlike local disk and Lustre, tmpfs exists within the page cache and should not 
be affected by it in the same way. However, when memory is low, tmpfs will 
begin to write to the swap space, located on disk. As a result, it is possible 
to see similar trends to local disk when space is low. 
Figure~\ref{fig:modeleval} (a) shows that the speed-up between in-memory 
computing and tmpfs increases as the D/C / $\delta$/$\gamma$ ratio increases. 
This is unexpected as both tmpfs and in-mem reside in memory and should 
therefore both have comparable makespans. This was found to occur when the 
contention to RAM was highest. In other words, when 750 chunks were written to
disk in 10 iterations with a task duration of 0.59s, Spark in-memory was 
approximately 2x faster. 
%why is this? will generate gantts to investigate further.

\subsection{Iterations}
\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{results/figures/iterations.pdf}
    \caption{Makespans of Spark and Nipype writing to memory, tmpfs, local 
             disk and Lustre for a given number of iterations. All 
             makespans use 125 chunks of the complete BigBrain and have a fixed
             task duration of 3.44 seconds}\label{fig:iterations}
\end{figure}

Figure~\ref{fig:iterations} shows the difference between the different 
filesystem choices given the number of iterations. At 1 iteration, all 
filesystems behave the same, as they are all writing to the page cache. As the 
total amount of data written by the application increases to 750~GB, there is a 
greater disparity between Lustre and in-memory (2.67~x slower, on average). 
Local disk performance, however, 
is still comparable to memory (1.38~x slower, on average). Despite local disk and 
Lustre both being in transient state, local disk encounters less contention 
than what would be found on Lustre.

%sp_mem 1it = 34s, 10it = 64 100it = 381
%3.44 * 1 = 3.44, 3.44*10 = 34, 3.44*100 = 344

%sp_tmpfs 1it = 31, 10it = 69,

At 100 iterations, or 7500~GB, Lustre can be found to be, on average, 3.82x 
slower than Spark in-memory. The page cache, on Zenith, occupies 20\% of 
available memory. Assuming all memory (187~GB) was available, only 37.4~GB of 
data could be held in a node's page cache at any given time. The slowdown 
experienced can be explained by the much less data could reside in the page 
cache at a given time, in contrast to 10 iterations, and therefore the effects 
of disk bandwidth are more significant in this workflow

While there is some variability that can be seen in Figure~\ref{fig:iterations} 
between the two engines, this believed to be insignificant, and potentially due 
to SLURM node allocation delays in our launching of Nipype.

% Comment on the general trend
% note 'on average' = average between spark and nipype. local disk spark is 
% 1.22x slower and local disk nipype is 1.54x slower
% spark lustre 2.66x slower; nipype lustre 2.68x slower

\subsection{CPU Time}
% Comment on the general trend
% We can't really say that imaging tasks are not small. Even Freesurfer could be decomposed.
%
\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{results/figures/cputime.pdf}
    \caption{Makespans of Spark and Nipype writing to memory, tmpfs, local 
             disk and Lustre with varying task durations. All 
             makespans use 125 chunks of the complete BigBrain and are fixed at 
             10 iterations}\label{fig:cputime}
\end{figure}

Increasing CPU time ensured that all filesystems had a comparable performance
(Figure~\ref{fig:cputime}). Lustre, for instance, is approximately 1.01x slower
than Spark in-memory at a task duration of 320 seconds, whereas it is 
approximately 3.25x slower that Spark in-memory with 2.4 second tasks. This 
pattern corroborated by our model which postulates that 
data movement costs will have little impact on compute-intensive tasks. The 
reasoning for this is that longer tasks give the page cache more time to flush 
between write to disk. The longer the delay, the less likely the page cache is 
to fill up when the concurrent data size written fits within the page cache.

\subsection{Image Block size}
\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{results/figures/numchunks.pdf}
    \caption{Makespans of Spark and Nipype writing to memory, tmpfs, local 
             disk and Lustre with varying numbers of chunks. All 
             makespans have the same overall sequential CPU time and 
             are fixed at 
             10 iterations}\label{fig:numchunks}
\end{figure}
% Comment on the general trend
As can be seen in Figure~\ref{fig:numchunks}, makespan decreases when the number 
of chunks increase. This is due to the fact that parallelism increases. With 
30 chunks, only 2 CPUs per node are actively working. At 125 chunks, this 
changes to a maximum of 9 CPUs per node, and at 750 chunks all 40 CPUs can be 
active. It can be noted that no significant speedup can be seen between 125 
chunks and 750 chunks. This is because out of the 50 chunks processed by each 
node, only 40 can be processed concurrently, thus requiring two sequential 
incrementation batches, and resulting in similar makespans as 125 
chunks.

In terms of filesystem selection, local and tmpfs perform comparably for all 
conditions, with Lustre being significantly slower. As with varying the number 
of iterations, Lustre is slower due to increased filesystem contention, which 
is, at minimum, 15x greater than contention on local disk. With an increase in 
number of chunks, local disk and tmpfs makespans begin to converge. A potential 
explanation for this may be that tmpfs is utilizing swap space. As concurrency 
increases, the memory footprint of the application also increases. It is 
possible that at 750 chunks, swapping to disk is required by tmpfs, thus 
resulting in similar processing times as local disk.

Swapping may also be an explanation for the variance between Spark in-memory 
and tmpfs performance. While Spark may also spill to disk, it only does so when
data does not fit in memory. As none of the RDDs generated throughout the 
pipeline were cached and all data concurrently accessed could be mantained 
in-memory, even for the 750 chunks which were processed in two batches, it did 
not need to spill to disk.

\subsection{Data Size}
\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{results/figures/datasize.pdf}
    \caption{Makespans of Spark and Nipype writing to memory, tmpfs, local 
             disk and Lustre with varying data sizes. All 
             makespans use 125 chunks and are fixed at 
             10 iterations with a task duration of 1.76 
             seconds }\label{fig:datasize}
\end{figure}
% Comment on the general trend
Increasing overall data size decreases performance, as can be seen in 
Figure~\ref{fig:datasize}. When the data size is very small, as is the case 
with the MRI image, all filesystem makespans are comparable. This is due to the 
fact that page cache can be leveraged fully regardless of file system. However, 
this time, Spark in-memory performed significantly worse than all other 
filesystems. Upon further inspection, it appeared that Spark in-memory executed
in a sequential order, on a single worker node. Lack of parallelism for the MRI 
image may be a result of Spark's max partition size, which is by default 128~MB
-- significantly larger than the 13~MB MRI image. Due to Spark in-memory's 
sequential execution, it was omitted from the graph.

At half BigBrain, the makespan differences become apparent in both local disk 
and Lustre, with Lustre, with Lustre becoming 2.4x slower than in-memory. This 
can be attributed to page cache saturation, as predicted by the model for both 
half the BigBrain image and the complete BigBrain. Only the MRI image was 
predicted to fall within the model constraints.

When the complete BigBrain is processed, the disparity between the different 
filesystems becomes even greater. Lustre becomes 3.68x slower, whereas local 
disk is now 1.68x slower. The reasoning for this is that the page cache fills 
up faster than with half of BigBrain.

\todo{change Spark's max partition size and test to see if it fixes the problem}
